Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-09-17 thesis/article.tex
Thesis: Updates Preliminaries, adds Pricing Games and Our model sections and improves previous ones

A \textit{flow} is a (stochastic) vector $x = (x_i)_{i \in N}$ where $x_i$ is the total flow on link $i$.
The flow then creates congestion on the links, each described by a non-decreasing latency function $(\ell_i)_{i \in N}$ which we assume to be affine.
We denote by $\mathcal{L}_d$ the class of polynomial latency functions with nonnegative coefficients and degree at most $d$, and as such $\mathcal{L}_1$ becomes the class of affine latency functions.
Finally, the network can also be extended by allowing a set of tolls $t = (t_i)_{i \in N}$ to be assigned to each link, which in turn adds to the effective cost of a player using them.

@@

In the homogeneous case all players have an equal reaction to tolls, and therefore the effective cost of a player using link $i$ becomes $\ell_i(x_i) + t_i$.
For a given set of tolls $t$, a flow $x$ is a \textit{Wardrop equilibrium for $t$} if $\forall i, j \in N$ with $x_i > 0$ it holds that
\begin{equation}
	\label{wardrop_inequality_cost}
	\ell_i(x_i) + t_i \leq \ell_j(x_j) + t_j
\end{equation}
In that case, all links with $x_i > 0$ have equal effective costs, i.e. there exists some $K > 0$ such that $\ell_i(x_1) + t_i = K$ for all links $i$ with $x_i > 0$.
From a well-known result from Beckman et al. \cite{beckmann1956studies} and Dafermos and Sparrow \cite{1363388843888284416}, such and an equilibrium for $t$ exists and is unique, therefore for a toll vector $t$ we can denote by $x(t)$ the flow occurring on  that unique equilibrium for the given toll $t$.
We copy Lemma 2.1 from Harkes et al \cite{Harks_2019} as is for later use.
\begin{lemma}
	\label{lemma:wardrop_equilibrium}
	A flow $x$ is a Wardrop equilibrium for $t$ if and only if for all feasible flows $x^\prime$,
	\[\sum_{i \in N} (\ell_i(x_i) + t_i) \cdot (x_i - x_i^\prime) \leq 0\]
\end{lemma}
If $t = 0$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at \eqref{wardrop_inequality_cost}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls.
I.e. if we consider any $c \in \R_+$ (or even $c \geq -\min_{i \in N} t_i$) then for $t^\prime = t + c$ the inequalities remain true for all $i, j \in N$ and thus the network game for the flow between tolls $t$ and $t^\prime$ is the same.

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(x, t) = \ell_i(x_i) + \alpha(p) t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Also we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Section ??, allowing many players close to $0$ breaks the pricing game.

@@

We introduce $n$ link owners which can assign a toll $t_i$ on link $i$ and profit from it.

\section*{2-link Pricing Games (our model)}

A 2-link $s-t$ parallel network is a graph with a source node $s$, a destination node $t$ and $2$ edges connecting $s$ to $t$.
Each link is associated with a latency function $l_1, l_2$
[TODO: Add figure]
A flow of infinitesimal users totaling up to a flow of $[0, 1]$ wishes to move from $s$ to $t$ through the link with the lowest cost.

@@

\begin{example}
	Consider a $2$-link parallel network pricing game with distribution function $\alpha$ for which it does not hold that $\alpha(p) > 0$ when $p > 0$.
\end{example}

>>>>

A \textit{flow} is a Lebesque measurable function $f: [0, 1] \rightarrow N$ that describes which link is selected by each player.
It is more intuitive, however, to consider the resulting flow on each link (\textit{flow on paths}), i.e. $x_i = \lambda(\{p \in [0, 1]: f(p) = i\})$.
Hence we get flow as a (stochastic) vector $x = (x_i)_{i \in N}$ where $x_i$ is the total flow on link $i$ with $x_i \geq 0$ and $\sum_{i \in N}x_i = 1$.
The flow then creates congestion on the links, each described by a non-decreasing latency function $(\ell_i)_{i \in N}$ which we assume to be affine.
We denote by $\mathcal{L}_d$ the class of polynomial latency functions with nonnegative coefficients and degree at most $d$, and as such $\mathcal{L}_1$ becomes the class of affine latency functions.
Finally, the network can also be extended by allowing a set of tolls $t = (t_i)_{i \in N}$ to be assigned to each link, which in turn adds to the effective cost of a player using them.

@@

In the homogeneous case all players have an equal reaction to tolls, and therefore the effective cost of a player using link $i$ becomes $\ell_i(x_i) + t_i$.
For a given set of tolls $t$, a flow $x$ is a \textit{Wardrop equilibrium for $t$} if $\forall i, j \in N$ with $x_i > 0$ it holds that
\begin{equation}
	\label{wardrop_inequality_cost}
	\ell_i(x_i) + t_i \leq \ell_j(x_j) + t_j
\end{equation}
In that case, all links with $x_i > 0$ have equal effective costs, i.e. there exists some $K > 0$ such that $\ell_i(x_1) + t_i = K$ for all links $i$ with $x_i > 0$.
From a well-known result from Beckman et al. \cite{beckmann1956studies} and Dafermos and Sparrow \cite{1363388843888284416}, such and an equilibrium for $t$ exists and is unique, therefore for a toll vector $t$ we can denote by $x(t)$ the flow occurring on  that unique equilibrium for the given toll $t$.
We will freely use when needed the general game-theoretical notation of $x(t_i, t_{-i})$ where $t_{-i} = t \setminus \{t_i\}$.
We copy Lemma 2.1 from Harkes et al \cite{Harks_2019} as is for later use.
\begin{lemma}
	\label{lemma:wardrop_equilibrium}
	A flow $x$ is a Wardrop equilibrium for $t$ if and only if for all feasible flows $x^\prime$,
	\[\sum_{i \in N} (\ell_i(x_i) + t_i) \cdot (x_i - x_i^\prime) \leq 0\]
\end{lemma}
If $t = 0$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at \eqref{wardrop_inequality_cost}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls.
I.e. if we consider any $c \in \R_+$ (or even $c \geq -\min_{i \in N} t_i$) then for $t^\prime = t + c$ the inequalities remain true for all $i, j \in N$ and thus the network game for the flow between tolls $t$ and $t^\prime$ is the same.

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(x(t)) = \ell_i(x_i) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Example \ref{example:a_0_0}, allowing many players close to $0$ breaks the pricing game.

With the above we can again define a Nash equilibrium for the game if $\forall i \in N$ and $\forall p \in [0, 1]$ it holds that
\[c_{f(p)}^p(x(t)) \leq c_i^p(x(t))\]
Existence of an equilibrium is guaranteed by the more general results of Schmeidler \cite[Thm 2]{1973JSP.....7..295S}, while uniqueness is covered by Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220}; for a more detailed coverage of those look at the definitions and propositions of Cole et al. \cite[\S2]{10.1145/780542.780618}.
Therefore can extend the definition for $x(t)$ similarly as in the homogeneous case.

@@

While keeping the above Network Game model, we introduce $n$ agents who each own one link of the network and can thus assign its toll and profit from it.
The profit is defined for each link owner $i$ as $\Pi_i = x_i \cdot t_i$ which, since we defined $x(t)$ above, we can extend the definition to $\Pi_i(t) = x_i(t) \cdot t_i$ describing the profit at the equilibrium for given toll $t$.
The link owners compete with each other for profit, each selfishly selecting the toll $t_i$ that will maximise their profit given the remaining tolls $t_{-i}$.
We describe this best response as $B_i(t_{-i}) = \argmax_{t_i \geq 0} \Pi_i(t_i, t_{-i})$.
Consequently, a toll vector $t$ is a \textit{Nash Equilibrium} for the pricing game, if $\forall i \in N$ and $\forall t_i^\prime \in \R$ it holds that
\[\Pi_i(t_i, t_{-i}) \geq \Pi_i(t_i^\prime, t_{-i})\]

For homogeneous players and affine latencies we can calculate the best response functions for each link owner by solving first order conditions on $\Pi_i(t_i, t_{-i})$ for $t_i$. TODO

\section*{2-link Heterogeneous Pricing Games (our model)}

Focusing on the model that will be used in this thesis, simplify our network by only having two links but consider that the flow is acting in a heterogeneous way.
The model can thus be defined as the tuple $(\ell, \alpha)$, where $\ell = (\ell_1, \ell_2)$ the latency functions and $\alpha$ the distribution function.
[TODO: Add figure]

@@

\begin{example}
	\label{example:a_0_0}
	Consider a $2$-link parallel network pricing game with distribution function $\alpha$ for which it does not hold that $\alpha(p) > 0$ when $p > 0$.
\end{example}
